Parsing Postfix Log Files

### 2009-08-27 doc/logparser-results.tex
Turned regex into an acronym everywhere.

The performance evaluation begins by describing the characteristics of the
mail server that produced the \numberOFlogFILES{} log files used to
evaluate the parser's performance and coverage; the computer that the tests
were run on is also described.  The characteristics of the
\numberOFlogFILES{} log files are described next, with an explanation of
why \parsername{} has better performance when parsing the larger log files
in the group.  The effect of using different rule orderings is explored,
and optimal rule ordering is compared to an oracle that allows the parser
to use only one rule when recognising each log line.  When \parsername{} is
used by other mail administrators, they will need to extend the ruleset to
parse their own log lines, so the next section addresses the question of
how parser performance is affected by an increase in the number of rules in
the ruleset.  The penultimate topic is the simple optimisation of caching
the results of compiling each rule's regex, and the huge effect it has on
parser efficiency.  The performance evaluation concludes by examining where
parsing time is spent: recognition of log lines or their subsequent
processing by actions?

@@

Before the Perl interpreter attempts to match a regex against a piece of
text, the regex is compiled into an internal representation and optimised
to improve the speed of matching.  This compilation and optimisation takes
CPU time: for most regexes it takes far more CPU time than the actual
matching does.  If the interpreter is certain that a regex will not change
while the program is running, it will automatically cache the results of
compiling and optimising the regex for later use.  The results of compiling
a dynamically generated regex can be cached and used in preference to the
original regex, but it is the responsibility of the programmer to do so;
\parsername{} does this with every rule's regex when the rules are loaded
from the database.

A test run using optimal ordering was performed as described in
\sectionref{parser efficiency}, with one difference: regexes were not
compiled and cached when the ruleset was loaded from the database, so the
Perl interpreter had to compile each regex each time it was used when
trying to recognise a log line.  \Graphref{Percentage increase in parsing
time when not caching compiled regexes graph} shows the effect that not
caching compiled regexes has on parser performance, with mean and standard
deviation in \tableref{Percentage increase in parsing time when not caching
compiled regexes table}.  For typical log files, the mean increase in
parsing time when not caching compiled regexes is
\input{build/include-cached-regexes-vs-discarded-regexes-mean}; looking at
it from the opposite direction, caching compiled regexes reduces parsing
time by
\input{build/include-cached-regexes-vs-discarded-regexes-mean-reduction}.
Caching compiled regexes is probably the single most effective optimisation
possible in \parsername{}, and was quite simple to implement: the framework
compiles each rule's regex when the ruleset is loaded, and uses the
compiled regex instead of the source regex when recognising log lines.  As
seen previously, log files 22 \& 62--68 do not suffer such a large increase
in parsing time when the optimisation is disabled; this is because, on
average, fewer rules are used per log line, so fewer regexes are compiled
per log line for those log files.

The increase in parsing time when parsing the first log file is much
greater than for the other log files (see \graphref{Percentage increase in
parsing time when not caching compiled regexes graph}); again, this is
because every rule's hits field starts at zero, so optimal ordering is less
efficient than usual, the mean number of rules used for each log line is
higher than usual, and more regexes will need to be compiled when
recognising each log line.  Parsing time is plotted against log file size
for caching and not caching compiled regexes in \graphref{Parsing time
plotted against log file size when caching and not caching compiled
regexes}; the plot for not caching compiled regexes is quite uneven
compared to caching compiled regexes --- some log files are particularly
slow, whereas others are not as badly affected, but these differences have
not been investigated.

@@

    \item Even with the optimisations described in this chapter enabled,
        recognising log lines still accounts for two thirds of the parser's
        execution time.  \Graphref{Percentage of parsing time spent
        recognising log lines graph} shows the percentage of parsing time
        spent recognising log lines for each of the \numberOFlogFILES{} log
        files, with mean and standard deviation shown in
        \tableref{Percentage of parsing time spent recognising log lines
        table}; for normal log files,
        \input{build/include-percentage-time-spent-recognising-log-lines-mean}
        of parsing time is spent recognising log lines.  To measure how
        long recognition of log lines takes, a full test run using optimal
        ordering and caching compiled regexes was performed as previously
        described in \sectionref{parser efficiency}, but with one
        difference: actions were not invoked when a log line was
        recognised, so the parsing time did not include the time normally
        taken by invocation of actions; the mean parsing times for each log
        file were subtracted from the mean parsing times from a normal test
        run to obtain the parsing time taken by actions.  The parsing times
        exclude the time taken for framework initialisation, loading and
        saving state tables, loading the ruleset, and other housekeeping
        tasks: in as far as possible, the times are just for recognising
        log lines or recognising log lines plus invoking actions.

        Optimal ordering reduces parsing time by
        \input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
        relative to shuffled ordering; actions occupy
        \input{build/include-percentage-time-spent-processing-log-lines-mean}
        of parsing time, less if any optimisations are disabled, so
        optimising actions could not possibly provide as big a performance
        increase as optimally ordering rules.  Similarly, caching compiled
        regexes provides an
        \input{build/include-cached-regexes-vs-discarded-regexes-mean-reduction}
        reduction in parsing time; not invoking actions at all reduces the
        most optimised parsing time by less than half that.  Optimising
        actions to 1\% of their original parsing time would be only
        slightly more effective than optimal ordering is, but would be
        vastly more difficult to implement; optimising actions would not
        reduce parsing time enough to justify the amount of effort
        required.

@@

Each Postfix component whose log lines are of interest must have at least
one rule that recognises its log lines, or all of its log lines will be
silently skipped; in the extreme case of an empty ruleset the parser would
skip every log line.  \parsername{} skips log lines from programs that do
not have any associated rules, because there may be any number of log lines
from other programs in the log file, and some Postfix components do not
produce any log lines of interest.  \parsername{} does not parse log lines
from non-Postfix programs, e.g.\ Amavisd-new or SpamAssassin; it could
easily be extended to do so, if a method could be developed to correctly
associate such log lines with existing state table entries.  To correctly
recognise all log lines that are not skipped, there must be a rule to
recognise each log line variant produced by each Postfix component; if a
log line is not recognised the parser will issue a warning, to inform the
user that they need to extend their ruleset.  Each rule's regex should be
as specific and precise as possible, to ensure accurate parsing: a rule
with a regex that matches zero or more of any character will recognise
every log line, but not in a meaningful way; most log lines contain fixed
strings, so this is not a problem in practice.

Full coverage of log lines can be achieved without undue effort, yet once
achieved it requires maintenance.  Maintaining full coverage is an ongoing
task because the set of log line variants changes over time, e.g.\
administrators add restrictions with custom messages, \acronym{DNSBL}
messages change, or major releases of Postfix change log lines (usually by
adding more information).  Warnings are issued for any log lines that are
not recognised; no warnings are issued for unrecognised log lines while
parsing the \numberOFlogFILES{} log files, so it can be safely concluded
that zero false negatives arise.  False positives are harder to quantify,
short of examining each of the 60,721,709 log lines and checking that the
correct rule recognised it.  However, a random sample of 6039 log lines was
parsed, and the results manually verified by inspection to ensure that the
correct rule recognised each log line.  The sample was generated by running
the command: \newline{} \tab{} \texttt{perl -n -e \singlequote{}print if
(rand 1 < 0.0001)\singlequote{} LOG\_FILES} \newline{} to randomly extract
roughly one log line in every 10,000 (it actually extracted 0.00994\%
instead of 0.010\%).  Each log line was examined and the correct rule
identified from the \numberOFrules{} rules in the database; the correct
rule was then compared to the rule that recognised the log line when
parsing.  The sample results contained zero false positives, and this check
has been automated to ensure continued accuracy.  Based on these results,
and how precise each rule's regex is, the author is confident that zero
false positives occur when parsing the \numberOFlogFILES{} log files.

@@

This chapter evaluated \parsername{} on two criteria: efficiency, and
coverage of Postfix log files.  The former began by describing the mail
server the log files were taken from, the computer used to generate the
statistics in this chapter, and the characteristics of the
\numberOFlogFILES{} log files used to generate statistics and test
coverage, including why performance is better when parsing the larger log
files.  The framework optimises the order in which rules are used when
trying to recognise each log line, and the effect that optimisation has on
parsing time is explored; this is followed by comparing optimal ordering
with an oracle.  How \parsername{} scales as the size of the ruleset
increases is addressed, followed by a description and analysis of the
simplest and most effective of the optimisations, caching compiled regexes,
and the efficiency evaluation concludes with an examination of where
parsing time is spent: recognising log lines or processing them?

>>>>

The performance evaluation begins by describing the characteristics of the
mail server that produced the \numberOFlogFILES{} log files used to
evaluate the parser's performance and coverage; the computer that the tests
were run on is also described.  The characteristics of the
\numberOFlogFILES{} log files are described next, with an explanation of
why \parsername{} has better performance when parsing the larger log files
in the group.  The effect of using different rule orderings is explored,
and optimal rule ordering is compared to an oracle that allows the parser
to use only one rule when recognising each log line.  When \parsername{} is
used by other mail administrators, they will need to extend the ruleset to
parse their own log lines, so the next section addresses the question of
how parser performance is affected by an increase in the number of rules in
the ruleset.  The penultimate topic is the simple optimisation of caching
the results of compiling each rule's \acronym{regex}, and the huge effect
it has on parser efficiency.  The performance evaluation concludes by
examining where parsing time is spent: recognition of log lines or their
subsequent processing by actions?

@@

Before the Perl interpreter attempts to match a \acronym{regex} against a
piece of text, the \acronym{regex} is compiled into an internal
representation and optimised to improve the speed of matching.  This
compilation and optimisation takes CPU time: for most \regexes{} it takes
far more CPU time than the actual matching does.  If the interpreter is
certain that a \acronym{regex} will not change while the program is
running, it will automatically cache the results of compiling and
optimising the \acronym{regex} for later use.  The results of compiling a
dynamically generated \acronym{regex} can be cached and used in preference
to the original \acronym{regex}, but it is the responsibility of the
programmer to do so; \parsername{} does this with every rule's
\acronym{regex} when the rules are loaded from the database.

A test run using optimal ordering was performed as described in
\sectionref{parser efficiency}, with one difference: \regexes{} were not
compiled and cached when the ruleset was loaded from the database, so the
Perl interpreter had to compile each \acronym{regex} each time it was used
when trying to recognise a log line.  \Graphref{Percentage increase in
parsing time when not caching compiled regexes graph} shows the effect that
not caching compiled \regexes{} has on parser performance, with mean and
standard deviation in \tableref{Percentage increase in parsing time when
not caching compiled regexes table}.  For typical log files, the mean
increase in parsing time when not caching compiled \regexes{} is
\input{build/include-cached-regexes-vs-discarded-regexes-mean}; looking at
it from the opposite direction, caching compiled \regexes{} reduces parsing
time by
\input{build/include-cached-regexes-vs-discarded-regexes-mean-reduction}.
Caching compiled \regexes{} is probably the single most effective
optimisation possible in \parsername{}, and was quite simple to implement:
the framework compiles each rule's \acronym{regex} when the ruleset is
loaded, and uses the compiled \acronym{regex} instead of the source
\acronym{regex} when recognising log lines.  As seen previously, log files
22 \& 62--68 do not suffer such a large increase in parsing time when the
optimisation is disabled; this is because, on average, fewer rules are used
per log line, so fewer \regexes{} are compiled per log line for those log
files.

The increase in parsing time when parsing the first log file is much
greater than for the other log files (see \graphref{Percentage increase in
parsing time when not caching compiled regexes graph}); again, this is
because every rule's hits field starts at zero, so optimal ordering is less
efficient than usual, the mean number of rules used for each log line is
higher than usual, and more \regexes{} will need to be compiled when
recognising each log line.  Parsing time is plotted against log file size
for caching and not caching compiled \regexes{} in \graphref{Parsing time
plotted against log file size when caching and not caching compiled
regexes}; the plot for not caching compiled \regexes{} is quite uneven
compared to caching compiled \regexes{} --- some log files are particularly
slow, whereas others are not as badly affected, but these differences have
not been investigated.

@@

    \item Even with the optimisations described in this chapter enabled,
        recognising log lines still accounts for two thirds of the parser's
        execution time.  \Graphref{Percentage of parsing time spent
        recognising log lines graph} shows the percentage of parsing time
        spent recognising log lines for each of the \numberOFlogFILES{} log
        files, with mean and standard deviation shown in
        \tableref{Percentage of parsing time spent recognising log lines
        table}; for normal log files,
        \input{build/include-percentage-time-spent-recognising-log-lines-mean}
        of parsing time is spent recognising log lines.  To measure how
        long recognition of log lines takes, a full test run using optimal
        ordering and caching compiled \regexes{} was performed as
        previously described in \sectionref{parser efficiency}, but with
        one difference: actions were not invoked when a log line was
        recognised, so the parsing time did not include the time normally
        taken by invocation of actions; the mean parsing times for each log
        file were subtracted from the mean parsing times from a normal test
        run to obtain the parsing time taken by actions.  The parsing times
        exclude the time taken for framework initialisation, loading and
        saving state tables, loading the ruleset, and other housekeeping
        tasks: in as far as possible, the times are just for recognising
        log lines or recognising log lines plus invoking actions.

        Optimal ordering reduces parsing time by
        \input{build/include-optimal-ordering-parsing-time-reduction-other-logs}\%
        relative to shuffled ordering; actions occupy
        \input{build/include-percentage-time-spent-processing-log-lines-mean}
        of parsing time, less if any optimisations are disabled, so
        optimising actions could not possibly provide as big a performance
        increase as optimally ordering rules.  Similarly, caching compiled
        \regexes{} provides an
        \input{build/include-cached-regexes-vs-discarded-regexes-mean-reduction}
        reduction in parsing time; not invoking actions at all reduces the
        most optimised parsing time by less than half that.  Optimising
        actions to 1\% of their original parsing time would be only
        slightly more effective than optimal ordering is, but would be
        vastly more difficult to implement; optimising actions would not
        reduce parsing time enough to justify the amount of effort
        required.

@@

Each Postfix component whose log lines are of interest must have at least
one rule that recognises its log lines, or all of its log lines will be
silently skipped; in the extreme case of an empty ruleset the parser would
skip every log line.  \parsername{} skips log lines from programs that do
not have any associated rules, because there may be any number of log lines
from other programs in the log file, and some Postfix components do not
produce any log lines of interest.  \parsername{} does not parse log lines
from non-Postfix programs, e.g.\ Amavisd-new or SpamAssassin; it could
easily be extended to do so, if a method could be developed to correctly
associate such log lines with existing state table entries.  To correctly
recognise all log lines that are not skipped, there must be a rule to
recognise each log line variant produced by each Postfix component; if a
log line is not recognised the parser will issue a warning, to inform the
user that they need to extend their ruleset.  Each rule's \acronym{regex}
should be as specific and precise as possible, to ensure accurate parsing:
a rule with a \acronym{regex} that matches zero or more of any character
will recognise every log line, but not in a meaningful way; most log lines
contain fixed strings, so this is not a problem in practice.

Full coverage of log lines can be achieved without undue effort, yet once
achieved it requires maintenance.  Maintaining full coverage is an ongoing
task because the set of log line variants changes over time, e.g.\
administrators add restrictions with custom messages, \acronym{DNSBL}
messages change, or major releases of Postfix change log lines (usually by
adding more information).  Warnings are issued for any log lines that are
not recognised; no warnings are issued for unrecognised log lines while
parsing the \numberOFlogFILES{} log files, so it can be safely concluded
that zero false negatives arise.  False positives are harder to quantify,
short of examining each of the 60,721,709 log lines and checking that the
correct rule recognised it.  However, a random sample of 6039 log lines was
parsed, and the results manually verified by inspection to ensure that the
correct rule recognised each log line.  The sample was generated by running
the command: \newline{} \tab{} \texttt{perl -n -e \singlequote{}print if
(rand 1 < 0.0001)\singlequote{} LOG\_FILES} \newline{} to randomly extract
roughly one log line in every 10,000 (it actually extracted 0.00994\%
instead of 0.010\%).  Each log line was examined and the correct rule
identified from the \numberOFrules{} rules in the database; the correct
rule was then compared to the rule that recognised the log line when
parsing.  The sample results contained zero false positives, and this check
has been automated to ensure continued accuracy.  Based on these results,
and how precise each rule's \acronym{regex} is, the author is confident
that zero false positives occur when parsing the \numberOFlogFILES{} log
files.

@@

This chapter evaluated \parsername{} on two criteria: efficiency, and
coverage of Postfix log files.  The former began by describing the mail
server the log files were taken from, the computer used to generate the
statistics in this chapter, and the characteristics of the
\numberOFlogFILES{} log files used to generate statistics and test
coverage, including why performance is better when parsing the larger log
files.  The framework optimises the order in which rules are used when
trying to recognise each log line, and the effect that optimisation has on
parsing time is explored; this is followed by comparing optimal ordering
with an oracle.  How \parsername{} scales as the size of the ruleset
increases is addressed, followed by a description and analysis of the
simplest and most effective of the optimisations, caching compiled
\regexes{}, and the efficiency evaluation concludes with an examination of
where parsing time is spent: recognising log lines or processing them?
